Wireless smart monitoring energy system: a proof of concept

### 2021-03-26 paper/main.tex
sicurezza

	\section{Related Work}
	Different designs have been proposed for this kind of application. In \cite{7884488} a wireless sensor network is implemented using ZigBee. In \cite{8612412}, \cite{8110428} the communication protocol used is TCP over WiFi. In our work we use UDP over the home WLAN. The theoretical advantage of using UDP is that we avoid the TCP overhead that is not needed in this kind of application, however, since the traffic generated is so little, empirical evidence of this advantage may only be manifest within a system with a large number of smart plugs connected. Differently from other  proposed works, on top of the energy monitoring system, we implement a mechanism that keeps the user from exceeding a configurable maximum power usage.

@@

	\section{Challenges}
	During the development, we found some issues outside the scope of this POC. Since the focus of the project was not on security we decided for ease of debugging to do not encrypt the packets exchanged between plugs and server. Another issue is that the configuration of the plug is hard-coded, for example, the Wi-Fi SSID and password are stored as a char array in our C++ code. Since the EPS 8266 could switch to access point mode which means generate a new wireless network, a configuration procedure could be implemented. A device - such as a smartphone - could connect to the plug wireless network and send the SSID and password required by our home wireless settings. That information should be stored in a persistent drive - like a micro SD card.
	As we already mentioned, another challenge to face is the security issue, the information sent from a plug to the server is completely clear, so there is no encryption implemented. A man-in-the-middle attack could intercept those packets and change their contents. An attacker could also pose as a smart plug, sending fake power readings to the server interfering with the normal operation.
	\subsection{Security}

>>>>

	\section{Related Work}
	Different designs have been proposed for this kind of application. In \cite{7884488} a wireless sensor network is implemented using ZigBee, which is more indicated for IoT scenarios. In \cite{8612412}, \cite{8110428} the communication protocol used is TCP over wireless. In our POC, we use UDP over a wireless connection. The theoretical advantage of using UDP is that we avoid the TCP overhead that is not needed in this kind of application, however, since the traffic generated is so little, empirical evidence of this advantage may only be manifest within a system with a large number of smart plugs connected. Differently from other proposed works, on top of the energy monitoring system, we implement a mechanism that keeps the user from exceeding a configurable maximum power usage.

@@

	\section{Challenges}
	During the development, we found some issues outside the scope of this POC. Since the focus of the project was not on security we decided for ease of debugging to do not encrypt the packets exchanged between plugs and server. Another issue is that the configuration of the plug is hard-coded, for example, the Wi-Fi SSID and password are stored as a char array in our C++ code. Since the EPS 8266 could switch to access point mode which means generate a new wireless network, a configuration procedure could be implemented. A device - such as a smartphone - could connect to the plug wireless network and send the SSID and password required by our home wireless settings. That information should be stored in a persistent drive - like a micro SD card.
	\section{Security issues}
	As we already mentioned, another challenge to face is the security issue, the information sent from a plug to the server is completely clear, so there is no encryption implemented. 
	In fact, our aim is to create a smart system and for debugging purpose is more convenient to use un-encrypted packets.
	 The following paragraphs describes different attacks which we have implemented or could be performed quite easily.
	\subsection{Fake plugs}
	A fake plug could be easily added to our system sending fake readings. In our scenario this could be a problem, because could lead to a different behavior.
	If an attacker uses a fake plug could send a random value to the system avoiding the user to turn on the correct device, considering that the system sums the energy consumption to prevent blackouts. An identification mechanism, for example using a token validation procedure, could be implemented to recognize and guarantee the access to authorized devices. An example of fake plug is the NodeMcu V3, which is used for debugging purpose, but from a different point of view could be see a malicious device and be a problem for a real scenario.
	\subsection{Man in the middle}
	A man in the middle attack could proxy the UDP traffic and change packets contents - such as the measurements or ON/OFF commands. 
	An attacker could even change the source IP address field of some INIT packets, and so the plugs would connect to the attacker server.
	Some tools, such as Scapy which is a Python library, allow to capture, modify and forward packets. Our system would not be able to detect that the packet has been modified, 
	also because the UDP checksum is not enabled and so, the integrity of header and payload is not guarantee.
